---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# contact

<!-- badges: start -->
<!-- badges: end -->

This package is designed to process spatiotemporal data into contact and social networks, and facilitate network analysis by randomizing individuals' movement paths and/or related categorical variables. To use this package, users need only have a dataset containing spatial data (i.e., latitude/longitude, or planar x & y coordinates), individual IDs relating spatial data to specific individuals, and date/time information relating spatial locations to temporal locations. The functionality of this package ranges from data "cleaning" via multiple filtration functions, to spatial and temporal data interpolation, and network creation and summarization. Functions within this package are not limited to describing interpersonal contacts. Package functions can also identify and quantify "contacts" between individuals and fixed areas (e.g., home ranges, water bodies, buildings, etc.). As such, this package is an incredibly useful resource for facilitating epidemiological, ecological, ethological and sociological research.

This package was created and is maintained by members of the [Lanzas Lab](http://www.lanzaslab.org/) at the North Carolina State University College of Veterinary Medicine.

## Installation

You can install the released version of contact from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("contact")

```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lanzaslab/contact")
```
## Example

*This example is an excerpt from the package vignette.*

Let's say you want to create a time-aggregated point-location-based environmental contact networ describing instances when animals, represented by data points, were in "contact" with a fixed location/area. As an example of this, here we use *contact* functions to describe contacts between tracked calves in a single feedlot pen and their water trough. 

```{r example}
library(contact)

data("calves") #load the calves data set

```

**A.) Ensure all required columns exist in the calves data set (i.e., xy coordinates, unique individual IDs, dateTime)**

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

head(calves)#The calves data set does not have a singular dateTime column. Rather, it has "date" and "time" columns. We must append a dateTime column to the data frame.

calves.dateTime<- contact::datetime.append(x = calves, date = calves$date, time= calves$time, dateTime = NULL, dateFormat = "mdy", dateFake = FALSE, startYear = NULL, tz.in = "UTC", tz.out = NULL, month = FALSE, day = FALSE, year = FALSE, hour = FALSE, minute = FALSE, second = FALSE, daySecond = FALSE, totalSecond = FALSE)

```

**B.) Calculate distances between the water-trough polygon and calves at each time step.**

First, we must define the location of the water trough. To do this, we read in point-location data for the water-trough vertices.

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

water<- data.frame(x = c(61.43315, 61.89377, 62.37518, 61.82622), y = c(62.44815, 62.73341, 61.93864, 61.67411)) #This is a data frame containing the x and y coordinates of the four trough vertices.

```

As noted in the dist2Area_df help documention, polygon-vertex coordinates must be arranged in a particular way. Here we arrange them accordingly.

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

water_poly<-data.frame(matrix(ncol = 8, nrow = 1)) #(ncol = number of vertices)*2
colnum = 0
for(h in 1:nrow(water)){
  water_poly[1,colnum + h] <- water$x[h] #pull the x location for each vertex
  water_poly[1, (colnum + 1 + h)] <- water$y[h] #pull the y location for each vertex
  colnum <- colnum + 1
}

```

Calculate distances between calves and the water polygon at every timestep.

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

water_distance<-contact::dist2Area_df(x = calves.dateTime, y = water_poly, x.id = "calftag", y.id = "water", dateTime = "dateTime", point.x = calves.dateTime$x, point.y = calves.dateTime$y, poly.xy = NULL, parallel = FALSE, dataType = "Point", lonlat = FALSE, numVertices = NULL) #note that the poly.xy and numVertices arguments refer to vertices of polygons in x, not y. Because dataType is "Point," not "Polygon," these arguments are irrelevant here.

head(water_distance)

```

**C.) Identify what SpTh value will allow us to capture 99% of contacts, defined as instances when point-locations were within 0.333 m of the water trough, given the RTLS accuracy.**

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

SpThValues<-contact::findDistThresh(n = 100000, acc.Dist1 = 0.5, acc.Dist2 = NULL, pWithin1 = 90, pWithin2 = NULL, spTh = 0.5) #spTh represents the initially-defined spatial threshold for contact. #spTh represents the initially-defined spatial threshold for contact. Note that we've chosen to use 100,000 in-contact point-location pairs here.

SpThValues #it looks like an adjusted SpTh value of approximately 0.71 m will likely capture 99% of contacts, defined as instances when point-locations were within 0.333 m of the water trough, given the RTLS accuracy. #Note that because these confidence intervals are obtained from distributions generated from random samples, every time this function is run, results will be slightly different. 

CI_99<-unname(SpThValues[21]) #we will use this SpTh value moving forward.

```

**D.) Identify time points when calves were within the re-adjusted SpTh distance from water trough.**

```{r, echo=TRUE, warning = FALSE, eval = TRUE}

water_contacts <- contact::contactDur.area(water_distance, dist.threshold=CI_99,sec.threshold=1, blocking = FALSE, equidistant.time = FALSE, parallel = FALSE, reportParameters = TRUE) #Note that because we are not interested in making a time-aggregated network with > 1 temporal levels, we set blocking = FALSE to reduce processing time.

head(water_contacts)

```


